---
title: "R-1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)

library(tidyverse)
library(plotly)
library(lubridate)
library(scales)
library(sf)
library(tigris)
library(leaflet)

options(tigris_use_cache = TRUE)
```

```{r county set-up}
ca_counties <- counties("CA", cb = T, progress_bar = F)
projection <- "+proj=utm +zone=10 +ellps=GRS80 +datum=NAD83 +units=ft +no_defs"

ca_counties_transformed <- 
  ca_counties %>% 
  st_transform(4326) %>% 
  st_transform(26910) %>% 
  st_transform(projection) %>% 
  st_transform(st_crs(ca_counties))

bay_county_names <-
  c("Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Francisco",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma")
bay_counties <-
  ca_counties %>%
  filter(NAME %in% bay_county_names)

usa_zips <- 
  zctas(cb = T, progress_bar = F)
bay_zips <-
  usa_zips %>% 
  st_centroid() %>% 
  .[bay_counties, ] %>% 
  st_set_geometry(NULL) %>% 
  left_join(usa_zips %>% select(GEOID10)) %>% 
  st_as_sf()
```

```{r file loading}

years <- 2017:2020
quarters <- 1:4
type <- "Electric"
pge_17_20_elec <- NULL

for (year in years) {
  for (quarter in quarters) {
    if (year < 2020) {
      filename <- paste0(
      "PGE_",year,"_Q",quarter,"_",type,"UsageByZip.csv"
        )
      temp <- read_csv(filename)
      pge_17_20_elec <- rbind(pge_17_20_elec, temp)
      saveRDS(pge_17_20_elec, "pge_17_20_elec.rds")
      }
    else {
      if (quarter < 3) {
        filename <- paste0(
        "PGE_",year,"_Q",quarter,"_",type,"UsageByZip.csv"
        )
        print(filename)
        temp <- read_csv(filename)
        pge_17_20_elec <- rbind(pge_17_20_elec, temp)
        saveRDS(pge_17_20_elec, "pge_17_20_elec.rds")
      }
    }
  }
}

type <- "Gas"
pge_17_20_gas <- NULL

for (year in years) {
  for (quarter in quarters) {
    if (year < 2020) {
      filename <- paste0(
      "PGE_",year,"_Q",quarter,"_",type,"UsageByZip.csv"
      )
      temp <- read_csv(filename)
      pge_17_20_gas <- rbind(pge_17_20_gas, temp)
      saveRDS(pge_17_20_gas, "pge_17_20_gas.rds")
      }
    else {
      if (quarter < 3) {
        filename <- paste0(
        "PGE_",year,"_Q",quarter,"_",type,"UsageByZip.csv"
        )
        temp <- read_csv(filename)
        pge_17_20_gas <- rbind(pge_17_20_gas, temp)
        saveRDS(pge_17_20_gas, "pge_17_20_gas.rds")
      }
    }
  }
}
```



```{r file filtering}

pge_17_20_elec <- mutate(pge_17_20_elec, TOTALKBTU = TOTALKWH*3412.14) 
pge_elec_filter <-
  pge_17_20_elec %>% 
  filter(CUSTOMERCLASS %in% c("Elec- Residential", "Elec- Commercial")) %>% 
  mutate(ZIPCODE = ZIPCODE %>% as.character()) %>%
  filter(ZIPCODE %in% bay_zips$ZCTA5CE10) %>% 
  group_by(ZIPCODE,CUSTOMERCLASS,MONTH,YEAR) %>% 
  summarize(TOTALKBTU = sum(TOTALKBTU, na.rm = T)) %>% 
  left_join(
    bay_zips %>% select(GEOID10),
    by = c("ZIPCODE" = "GEOID10")
    ) %>% 
  st_as_sf() %>% 
  st_transform(4326)
  

pge_17_20_gas <- mutate(pge_17_20_gas, TOTALKBTU = TOTALTHM*99976.1)

pge_gas_filter <-
  pge_17_20_gas %>%
  filter(CUSTOMERCLASS %in% c("Gas- Residential", "Gas- Commercial")) %>%
  mutate(ZIPCODE = ZIPCODE %>% as.character()) %>% 
  filter(ZIPCODE %in% bay_zips$ZCTA5CE10) %>%
  group_by(ZIPCODE,CUSTOMERCLASS,MONTH,YEAR) %>%
  summarize(TOTALKBTU = sum(TOTALKBTU, na.rm = T)) %>% 
  left_join(
    bay_zips %>% select(GEOID10),
    by = c("ZIPCODE" = "GEOID10")
  ) %>%
  st_as_sf() %>%
  st_transform(4326)

pge_final <- rbind(pge_elec_filter,pge_gas_filter)
pge_final$DATE <- paste(pge_final$YEAR, "-", pge_final$MONTH, "-01", sep="") %>% as.Date(format = "%Y-%m-%d")
```


```{r graph}
pge_chart <-
  pge_final %>% 
  ggplot() +
  geom_bar(
    aes(
      x = DATE,
      y = TOTALKBTU,
      fill = CUSTOMERCLASS
    ),
    stat = "identity",
    position = "stack"
  ) + 
  labs(
    x = "Month",
    y = "kBTU",
    title = "PG&E Bay Area Monthly kBTU Usage, 2017-2020(Q1+Q2)",
    fill = "Electricity Type"
  ) + scale_x_date(date_breaks = "1 month", date_labels ="%m-%y", expand = c(0,0)) +  
      theme(axis.text.x = element_text(size = 6, angle = 45, colour = "black",
      vjust = 1, hjust = 1))
pge_chart
```
```{r}

pge_19_elec <- pge_17_20_elec %>% filter(YEAR == "2019") %>% 
  filter(MONTH == 3) %>% 
  filter(CUSTOMERCLASS %in% c("Elec- Commercial","Elec- Residential")) %>% 
  mutate(ZIPCODE = ZIPCODE %>% as.character()) %>%
  filter(ZIPCODE %in% bay_zips$ZCTA5CE10) %>% 
  group_by(ZIPCODE) %>% 
  summarize(TOTALKWH = sum(TOTALKWH, na.rm = T)) %>% 
  left_join(
    bay_zips %>% select(GEOID10),
    by = c("ZIPCODE" = "GEOID10")
  ) %>% 
  st_as_sf() %>% 
  st_transform(4326)

pge_20_elec <- pge_17_20_elec %>% filter(YEAR == "2020") %>% 
  filter(MONTH == 3) %>% 
  filter(CUSTOMERCLASS %in% c("Elec- Commercial","Elec- Residential")) %>% 
  mutate(ZIPCODE = ZIPCODE %>% as.character()) %>%
  filter(ZIPCODE %in% bay_zips$ZCTA5CE10) %>% 
  group_by(ZIPCODE) %>% 
  summarize(TOTALKWH = sum(TOTALKWH, na.rm = T)) %>% 
  left_join(
    bay_zips %>% select(GEOID10),
    by = c("ZIPCODE" = "GEOID10")
  ) %>% 
  st_as_sf() %>% 
  st_transform(4326)

pge_1920_elec <- cbind(pge_20_elec,pge_19_elec)
pge_1920_elec <- mutate(pge_1920_elec, CHANGE = TOTALKWH-TOTALKWH.1)
```
From the previous graph, it appears that electricity consumption in the Bay Area is slightly greater in March 2020 (shelter-in-place orders were first enacted on March 16) than in March 2019. While sheltering in place may lend itself to increased electricity consumption residentially rather than commercially, it is unclear how that manifests on the following graph, as changes do not seem to correspond to urban centers. Other factors, such as weather, may influence regional changes in electricity consumption.
```{r map}

res_pal <- colorNumeric(
  palette = "viridis",
  domain = 
    pge_1920_elec$CHANGE
)

leaflet() %>% 
  addTiles() %>% 
  addPolygons(
    data = pge_1920_elec,
    fillColor = ~res_pal(CHANGE),
    color = "white",
    opacity = 0.5,
    fillOpacity = 0.5,
    weight = 1,
    label = ~paste0(
      round(CHANGE), 
      " kWh change in ",
      ZIPCODE
    ),
    highlightOptions = highlightOptions(
      weight = 2,
      opacity = 1
    )
  ) %>% 
  addLegend(
    data = pge_1920_elec,
    pal = res_pal,
    values = ~CHANGE,
    title = "Regional Change in kWh, comparing March of 2019 and 2020"
  )
```